---
title: "CKME136"
author: "Hira Khan"
date: '2018-12-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(mapproj)
library(noncensus)
library(data.table)
library(corrplot)
library(fiftystater)
dataset<-read.csv(file.choose(),header = TRUE,stringsAsFactors = FALSE)
#dataset<-read.csv("healthsearch2.csv",header = TRUE, stringsAsFactors = FALSE)

```

Cleaning Dataset
```{r}
names(dataset)[1] <- "USA" # Rename the first column to 'USA'
dataset_1 <- dataset[1:2]
dataset_2 <- dataset[3:length(dataset)]
names(dataset_2) <- str_sub(names(dataset_2), 2, -1) # Use str_sub to get the year.condition values
dataset_1[which(dataset_1$USA == "Washington DC (Hagerstown MD)"), 1] <- "Washington DC (Hagerstown) MD"

```

Load the states dataset to join to healthsearch dataset
```{r}
data(states)
#join year.condtion column and then separate into two separate columns 
dd = dplyr::bind_cols(dataset_1, dataset_2)
dataset <- dd %>%
  gather("year.condition", "searches", 3:length(dataset))  %>%
  separate(col = "year.condition", into = c("year", "condition"), sep = "\\.") %>%
  separate(col = "USA", into = c("city", "state"), sep = -3) %>%   #Changed -2 to -3
  group_by(city, state, geoCode, year, condition) %>%
  summarize(searches = sum(searches)) %>%
  left_join(states, by = "state") %>% # Join states table to get full names and population
  select(3, 1:2, 7:10, 4:6, 11:12) %>%
  setnames(old = c("state", "name"), new = c("state_abb", "state_name"))

head(dataset, 10)

```

PLOTS
```{r}
searches_by_year <- dataset %>% 
  group_by(year) %>%
  summarize(searches = sum(searches)) %>%
  mutate(difference = searches - lag(searches, default = first(searches)))
ggplot(searches_by_year, aes(x = year, y = searches, group = 1)) +
  geom_line(col = "darkblue", size = 2) +
  labs(title = "Health Search Volume by 2004-2017")

```

PLOT for searches by year and condition
```{r}
searches_by_condition <- dataset %>%
  group_by(year, condition) %>%
  summarize(searches = sum(searches))
ggplot(searches_by_condition, aes(x = year, y = searches, group = condition, col = condition)) +
  geom_smooth(size = 1, se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(color = "black")) +
  labs(title = "Health Search Volume by Year and Condition", subtitle = "2004 - 2017") 
```

Correlation between 9 condtions (using tidyr,dplyr,corrplot packages )
```{r}
dataset <- ungroup(dataset)
yc_spread <- dataset %>%
  spread(key = condition, value = searches)
yc_cor <- yc_spread %>%
  select(11:19) 
condition_cor <- cor(yc_cor, method = "pearson")
corrplot(condition_cor, method = "number", type = "upper", order = "hclust", tl.col = "red",tl.srt = 40)
corrplot.mixed(condition_cor,lower.col = "black", number.cex=.80)
```


Data Preprocessing 
```{r}
#rename yc_spread to healthsearch 
healthsearch <-yc_spread[1:length(yc_spread)]
healthsearch$state_name<-NULL
healthsearch$region<-NULL
healthsearch$capital<-NULL
healthsearch$area<-NULL
healthsearch$population<-NULL
healthsearch$division<-NULL
head(healthsearch)

basics <- function(att) {
  cat(att, "\n",
      "max:", max(healthsearch[[att]]), "\n",
      "min:", min(healthsearch[[att]]), "\n",
      "mean:", mean(healthsearch[[att]]), "\n",
      "sd:", sd(healthsearch[[att]]), "\n\n"
      );
  }

#calling the above function for numerical attributes 
basics("cancer");
basics("cardiovascular");
basics("depression");
basics("diabetes")
basics("diarrhea")
basics("obesity")
basics("rehab")
basics("stroke")
basics("vaccine")

#will disregard geocode, since it will not help in predictive analysis
healthsearch$geoCode<-NULL

```

PLOTS for numerical attributes (all conditions)
```{r}
hist(healthsearch$cancer, breaks = 20)
hist(healthsearch$cardiovascular, breaks = 20)
hist(healthsearch$depression, breaks = 20)
hist(healthsearch$diabetes, breaks = 20)
hist(healthsearch$diarrhea, breaks = 20)
hist(healthsearch$obesity, breaks = 20)
hist(healthsearch$rehab, breaks = 20)
hist(healthsearch$stroke, breaks = 20)
hist(healthsearch$vaccine, breaks = 20)

#cancer-normal distribution
#cardiovascular-right skewed
#all histograms have significant outliers 

```

DETERMINE OUTLIER VALUES

Using interquartile range method
(Outliers are more than 1.5xIQR below Q1 or 1.5xIQR above Q3):
```{r}
upperWhisker <- function(att) {
  Q1 <- as.numeric(quantile(healthsearch[[att]], c(0.25)));
  Q3 <- as.numeric(quantile(healthsearch[[att]], c(0.75)));
  return (Q3 + 1.5*(Q3 - Q1));
}

lowerWhisker <- function(att) {
  Q1 <- as.numeric(quantile(healthsearch[[att]], c(0.25)));
  Q3 <- as.numeric(quantile(healthsearch[[att]], c(0.75)));
  return (Q1 - 1.5*(Q3 - Q1));
}

```
Using standard deviation method
(Outliers are more than 3 standard deviations from the mean):

```{r}
upper3SD <- function(att) {
  return (mean(healthsearch[[att]]) + 3 * sd(healthsearch[[att]]));
}

lower3SD <- function(att) {
  return (mean(healthsearch[[att]]) - 3 * sd(healthsearch[[att]]));
}
```

Set upper and lower thresholds to eliminate values that meet at least one of the above criteria:

```{r}
upperOutlier <- function(att) {
  return (min(max(upper3SD(att), upperWhisker(att)), max(healthsearch[[att]])))
}

lowerOutlier <- function(att) {
  return (max(min(lower3SD(att), lowerWhisker(att)), min(healthsearch[[att]])))
}
```
Here is the proportion of outliers for each numerical variable:

```{r}
outlierProp <- function(att) {
  numAll <- length(healthsearch[[att]]);
  numOutliers <- length(healthsearch[[att]][(dataset[[att]] > upperOutlier(att)) | (healthsearch[[att]] < lowerOutlier(att))]);
  cat(att, ":", (numOutliers / numAll), "\n")
  }
outlierProp("cancer")
outlierProp("cardiovascular")
outlierProp("depression")
outlierProp("diabetes")
outlierProp("diarrhea")
outlierProp("obesity")
outlierProp("rehab")
outlierProp("stroke")
outlierProp("vaccine")

#there were no outliers in the dataset 
hist(healthsearch$cancer, breaks = 20)
hist(healthsearch$cardiovascular, breaks = 20)
hist(healthsearch$depression, breaks = 20)
hist(healthsearch$diabetes, breaks = 20)
hist(healthsearch$diarrhea, breaks = 20)
hist(healthsearch$obesity, breaks = 20)
hist(healthsearch$rehab, breaks = 20)
hist(healthsearch$stroke, breaks = 20)
hist(healthsearch$vaccine, breaks = 20)

```

CORRELATIONS

Pearson correlation of numerical variables covering entire dataset
(+1 = strong positive correlation
  0 = no correlation 
 -1 = strong negative correlation)

```{r}
cordataset <- subset(healthsearch, select = c("cancer", "cardiovascular", "depression", "diabetes","diarrhea","obesity","rehab","stroke","vaccine"))
cor(cordataset)

#all attributes have a strong positive correlation 
```

NORMALIZATION

This adjusts for any bias due to the large numerical values associated with some attributes.

```{r}
healthsearch$cancer <- scale(healthsearch$cancer)
healthsearch$cardiovascular <- scale(healthsearch$cardiovascular)
healthsearch$depression <- scale(healthsearch$depression)
healthsearch$diabetes <- scale(healthsearch$diabetes)
healthsearch$diarrhea <- scale(healthsearch$diarrhea)
healthsearch$obesity <- scale(healthsearch$obesity)
healthsearch$rehab <- scale(healthsearch$rehab)
healthsearch$stroke <- scale(healthsearch$stroke)
healthsearch$vaccine <- scale(healthsearch$vaccine)
```
